Intrinsic alignments from projected tidal fields: non-linear impact on cosmic shear probes

### 2024-02-19 paper_IA.tex
Added first MCMC chains

Our infusion method completely relies on couplings between intrinsic galaxy shapes and the local tidal field, hence the first step in our method consist in extracting the tidal field maps $s_{ij}(\theta,\phi,\chi_i)$ from the mass maps $\delta(\theta,\phi,\chi_i)$ that source them. In three dimensions, trace-free tidal tensor $s_{ij}(\boldsymbol x)$ can be obtained from the matter over-density field $\delta(\boldsymbol x)$ as   \citep{Catelan_IA_Tidal}:
\begin{eqnarray}
 \widetilde{s}_{ij} (\boldsymbol k)  = \left[\frac{k_i k_j}{k^2} - \frac{\delta_{ij}}{3}\right]  \widetilde {\delta}(\boldsymbol k) \mathcal{G}(\sigma_{\rm G}) \, ,
 \label{eq:sij}
\end{eqnarray}
where $\mathcal{G}(\sigma_{\rm G})$ is a three-dimensional Gaussian function described by a single (free) parameter $\sigma_{\rm G}$ that  controls the physical scales which are allowed to affect the IA term in our model. 
Tilde symbols  denote Fourier transformed quantities,  the indices $(i,j)$ label the components of the Cartesian wave-vector $\boldsymbol{k}^T = (k_x,k_y,k_z)$, and $k^2 = k_x^2 + k_y^2 + k_z^2$. 
As shown in \citet{Tidalator} for Cartesian coordinates, projected tidal fields  computed from projected mass sheets provide and excellent agreement with the NLA model, which in contrast computes the full tidal fields from the three-dimensional matter density and project along the radial dimension at the end. We promote this transformation to curved-sky in this work, exploiting the  polarisation {\sc alm2map} operations built in {\sc Healpy}. In particular, noting that $Q=s_{xx}-s_{yy}$, $U=s_{xy}$ and $\delta=s_{xx}+s_{yy}$, we compute the curved-sky tidal field tensors ${s}_{ij}({\boldsymbol \theta})$ as:
 \begin{eqnarray}
s_{xx}({\boldsymbol \theta})  =  \left[\frac{ \delta + Q}{2}  - \frac{\delta}{3}\right] \, ,s_{yy}({\boldsymbol \theta})  =  \left[\frac{\delta - Q}{2}  - \frac{\delta}{3}\right] \, , s_{xy}({\boldsymbol \theta}) = U
 \label{eq:sij_2D_sph}
 \end{eqnarray}
where the $U({\boldsymbol \theta})$ and $Q({\boldsymbol \theta})$ maps are smoothed by the Gaussian beam with width $\sigma_{\rm G}$. We suppress large artificial tidal fields at the maps boundary by replicating 8$\times$ the density maps defined on the octant and working on a full sky density $\delta$; we re-apply the octant mask on the tidal field maps after the operation. Note that the value of $\sigma_{\rm G}$ is a free parameter both in the infusion technique described in this paper and in the NLA and TATT models. We therefore explore two cases, $\sigma_{\rm G}=0.1$Mpc and 0.5Mpc, however this may be further optimised in the future.

@@

%\subsection{Validation with full inference}
 %\label{sec:inference}

%priors, firecrown, Takahashi $P(k)$

@@

>>>>

Our infusion method completely relies on couplings between intrinsic galaxy shapes and the local tidal field, hence the first step in our method consist in extracting the tidal field maps $s_{ij}(\theta,\phi,\chi_i)$ from the mass maps $\delta(\theta,\phi,\chi_i)$ that source them. In three dimensions, trace-free tidal tensor $s_{ij}(\boldsymbol x)$ can be obtained from the matter over-density field $\delta(\boldsymbol x)$ as   \citep{Catelan_IA_Tidal}:
\begin{eqnarray}
 \widetilde{s}_{ij} (\boldsymbol k)  = \left[\frac{k_i k_j}{k^2} - \frac{\delta_{ij}}{3}\right]  \widetilde {\delta}(\boldsymbol k) \mathcal{G}(\sigma_{\rm G}) \, ,
 \label{eq:sij}
\end{eqnarray}
where $\mathcal{G}(\sigma_{\rm G})$ is a three-dimensional Gaussian function described by a single (free) parameter $\sigma_{\rm G}$ that  controls the physical scales which are allowed to affect the IA term in our model. 
Tilde symbols  denote Fourier transformed quantities,  the indices $(i,j)$ label the components of the Cartesian wave-vector $\boldsymbol{k}^T = (k_x,k_y,k_z)$, and $k^2 = k_x^2 + k_y^2 + k_z^2$. 
As shown in \citet{Tidalator} for Cartesian coordinates, projected tidal fields  computed from projected mass sheets provide and excellent agreement with the NLA model, which in contrast computes the full tidal fields from the three-dimensional matter density and project along the radial dimension at the end. We promote this transformation to curved-sky in this work, exploiting the  polarisation {\sc alm2map} operations built in {\sc Healpy}. In particular, noting that $Q=s_{xx}-s_{yy}$, $U=s_{xy}$ and $\delta=s_{xx}+s_{yy}$, we compute the curved-sky tidal field tensors ${s}_{ij}({\boldsymbol \theta})$ as:
 \begin{eqnarray}
s_{xx}({\boldsymbol \theta})  =  \left[\frac{ \delta + Q}{2}  - \frac{\delta}{3}\right] \, ,s_{yy}({\boldsymbol \theta})  =  \left[\frac{\delta - Q}{2}  - \frac{\delta}{3}\right] \, , s_{xy}({\boldsymbol \theta}) = U
 \label{eq:sij_2D_sph}
 \end{eqnarray}
where the $U({\boldsymbol \theta})$ and $Q({\boldsymbol \theta})$ maps are smoothed by the Gaussian beam with width $\sigma_{\rm G}$. We suppress large artificial tidal fields at the maps boundary by replicating 8$\times$ the density maps defined on the octant and working on a full sky density $\delta$; we re-apply the octant mask on the tidal field maps after the operation. Note that the value of $\sigma_{\rm G}$ is a free parameter both in the infusion technique described in this paper and in the NLA and TATT models. We therefore explore two cases, $\sigma_{\rm G}=0.1$Mpc and 0.5Mpc, however this may be further optimised in the future. Finally, the full-sky density field is downgraded from $N_{\rm side}=8192$ to $N_{\rm side}=4096$ since the smoothing washed the smallest angular scales.

@@

\subsection{Validation with full inference}
 \label{sec:inference}
\begin{figure}
%\includegraphics[width=\columnwidth]{graphs/triangle_nla_5D.pdf}
\includegraphics[width=\columnwidth]{graphs/triangle_bta_sweep.pdf}
\caption{Cosmological inference from the NLA-infused simulations, analysed the extended-NLA model. The cross-hairs represent the truth as fixed in the OuterRim $N$-body simulations \citep{OuterRim}.}
\label{fig:triangle_nla}
\end{figure}
priors, {\sc CosmoSIS}, Takahashi $P(k)$, etc.

@@

\begin{figure*}
\includegraphics[width=\columnwidth]{graphs/xip_sims_vs_th.eps}
\includegraphics[width=\columnwidth]{graphs/xim_sims_vs_th.eps}
\caption{Fractional difference on $\xi_\pm$ between the measurements from SkySim5000 and the model predictions.  }
\label{fig:frac_err_sims_th}
\end{figure*}
